%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
\begin{center}
\textit{Gratitude makes sense of our past,\\ brings peace for today,\\ and creates a vision for tomorrow.}\\
\textbf{--Melody Beattie}
\end{center}

\vspace{1cm}

Completing this thesis has been a challenging yet rewarding journey, and I am deeply grateful to all those who have supported me throughout this process.

First and foremost, I would like to express my deepest gratitude to my supervisor, \textit{Prof. Stephan Fritzsche}, for accepting me as his PhD student and providing me with the extraordinary opportunity to grow as a researcher. His trust in my potential from the very beginning gave me the confidence to pursue challenging questions and push the boundaries of my understanding.

I am grateful to \textit{Prof. Fritzsche} for his exceptional mentorship, which extended far beyond the boundaries of academic supervision. His keen eye for scientific writing not only refined my technical expression but also transformed the way I communicate research, teaching me how to craft a compelling narrative that resonates with the scientific community. I am also grateful to \textit{Dr. Bindiya Arora} for guiding me throughout my research journey. From my bachelor’s to my master’s degree, she introduced me to the exciting world of science and gave me my first real experience in research. Her constant support and belief in me helped me become the researcher I am today. This thesis is a reflection of everything she taught me from the very beginning.

I am also grateful to \textit{Dr. Birger Böning} for first introducing me to the fascinating field of strong-field physics. His enthusiasm and expertise ignited my curiosity and laid the foundation for my academic journey. Without his early mentorship, introducing me to key concepts, sharing his profound insights, and encouraging me to explore bold ideas, this thesis would not have taken shape. Dr. Böning’s generosity with his time and knowledge opened doors I had not imagined, and for that, I will always be grateful.

I am equally grateful to all my colleagues for stimulating discussions, collaborative meetings, and the vibrant academic environment we shared. Our exchanges of ideas, whether during coffee breaks or seminars fundamentally shaped my research perspective. \textit{Dr. Giorgio Visentin} , my officemate, thank you for all the laughs, the random chats about life, and the moments when we both just needed a break from science. You made the PhD grind so much lighter, and I’ll always remember our shared laughs at whatever nonsense the day brought. This journey would’ve been way less fun without you.

A special thanks goes to \textit{Dr. Shreyas Ramakrishna}, \textit{Dr. Aloka Kumar Sahoo} and \textit{Richard Bernecker} for generously dedicating their time to proofread my thesis. Their sharp corrections and insightful comments on technical phrasing greatly enhanced the quality of my thesis. Additionally, I deeply appreciate \textit{Richard Bernecker} for his careful review of the abstract and his precise corrections to the German translation.

I would also like to express my heartfelt thanks to my friends, especially \textit{Dr. Fang Liu} and \textit{Joseph Andrews}, for making my time in Jena truly unforgettable. Their companionship, support, and countless shared moments turned challenges into adventures and made this journey all the more meaningful. Beyond academic pursuits, they helped create a home away from home. I must specially thank \textit{Sumeet}, who was my first point of contact in Germany. From helping me navigate the initial culture shock to being a constant friend through visits to Erlangen, her support formed the foundation of my positive experience here.

Above all, I owe my deepest gratitude to my family, especially my parents, for their unconditional love and endless support throughout this journey. Their sacrifices paved the way for my education, their encouragement carried me through challenges, and their belief in me never wavered, even across continents. This achievement belongs as much to them as it does to me.



\endgroup
